---
output:
  pdf_document: default
header-includes:
  - \usepackage{color}
  - \usepackage[dvipsnames]{xcolor}
---
----
 Fall 2019: MATH 347 Bayesian Statistics
---

## \textcolor{RoyalBlue}{R Script for the All-nighters Problem}

Here we use R for the Bayesian posterior distribution with Binomial data and a discrete prior. A script to estimate the posterior distribution has been created and provided here. 
You can copy this directly into R and it will run all the commands. 


#### \textcolor{RoyalBlue}{Step 1: Load the prior and data.}

```{r}
priorvalues <- c(0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1)
priorprob <- c(1/23, 1/23, 7/23, 7/23, 3/23, 3/23, 1/23, 0/23, 0/23, 0/23, 0/23)

n <- 10
y <- 3
```

#### \textcolor{RoyalBlue}{Step 2: Calculating the likelihood probabilities and posterior probabilities.}
```{r}
#vector for storing results
jointprob <- numeric(length = length(priorvalues))

for(i in 1:length(priorvalues))
{
  
  #compute Binomial probability given value of p - likelihood
  binomprob <- dbinom(y, n, p = priorvalues[i])
  
  #compute joint probability - posterior
  jointprob[i] <- binomprob * priorprob[i]
  
}

#compute marginal probability of y 
pofy <- sum(jointprob)

#compute posterior probabilities
posteriorprob <- jointprob/pofy
```

#### \textcolor{RoyalBlue}{Step 3: Present and visualize results.}
```{r}
#put posterior probabilities in one matrix object for easy viewing 
allnighterposterior <- as.data.frame(cbind(priorvalues, priorprob, posteriorprob))
names(allnighterposterior) <- c("p", "prior", "posterior")

#list the final posterior distribution, based on our prior derived in class
allnighterposterior

#plot the prior and posterior probabilities
require(ggplot2)
require(reshape2)

allnighterposterior_all <- melt(allnighterposterior, id = "p")

ggplot(allnighterposterior_all, aes(x = p, y = value, colour = variable)) +
  geom_point(size = 3) + 
  xlab("p") + ylab("probability") +
 theme_bw(base_size = 12, base_family = "")

```
