---
title: Bayesian Inference for a Proportion
author: Jingchen (Monika) Hu 
institute: Vassar College
date: MATH 347 Bayesian Statistics
output:
  beamer_presentation:
    includes:
      in_header: ../LectureStyle.tex
slide_level: 2
fontsize: 11pt

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
require(gridExtra)
require(ProbBayes)
require(tidyverse)
crcblue <- "#2905a1"
knitr::opts_chunk$set(echo = TRUE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

## Outline

\tableofcontents[hideallsubsections]

# Example: Tokyo Express customers' dining preference

## Tokyo Express customers' dining preference
- Tokyo Express is a popular Japanese restaurant on Lagrange Ave (10-15 minutes walk from Vassar).
\pause
- Suppose the restaurant owner wants to improve the business even more, especially for dinner.
\pause
- The owner plans to conduct a survey by asking their customers: ``what is your favorite day to eat out for dinner?"
- The owner wants to find out how popular is choice of $\color{VassarRed}{\text{Friday}}$.


# Bayesian inference with discrete priors

## Bayesian inference with discrete priors
- Three general steps of Bayesian inference:
    - Step 1: express an opinion about the location of the proportion $p$ before sampling (prior).
    - Step 2: take the sample and record the observed proportion of preferring Friday (data/likelihood).
    - Step 3: use Bayes' rule to sharpen and update the previous opinion about $p$ given the information from the sample (posterior).
\pause

- Bayesian inference with $\color{VassarRed}{\text{discrete priors}}$:
    - Step 1 on priors: list the $\color{VassarRed}{\text{finite number}}$ of possible values for the proportion $p$, and assign probability to each value.
    \pause
    - Step 2 on data/likelihood: $\color{VassarRed}{\text{Binomial distribution}}$.
    \pause
    - Step 3 on posterior: use the discrete version of the $\color{VassarRed}{\text{Bayes' rule}}$ (summation $\Sigma$) for to sharpen and update the probability of each specified possible values of $p$.


## Step 1: Prior distribution
- Consider the percentage of customers' choice is Friday, $p$.
- $\color{VassarRed}{\text{Before giving out the survey}}$, let's consider:
    - the possible value(s) of proportion $p$;
    - the probability associate with each value of $p$.
\pause
- Suppose $p$ can take 6 possible values
  \begin{eqnarray}
  p = \{0.3, 0.4, 0.5, 0.6, 0.7, 0.8\}.
  \label{eq:dining:p}
  \end{eqnarray}
\pause
- Further suppose Tokyo Express owner believes that some values are more likely than the
others, specifically, a prior distribution:
  \begin{eqnarray}
  \pi_{owner}(p)= (0.125, 0.125, 0.250, 0.250, 0.125, 0.125).
  \label{eq:dining:expertprior}
  \end{eqnarray}
- Exercise: Is the prior distribution $\pi_{owner}(p)$ reasonable? (Hint: 3 axioms of probability)


## Using R/RStudio to express and plot the prior $\pi_{owner}(p)$

```{r fig.height = 2, fig.width = 2, fig.align = "center", size = "footnotesize"}
bayes_table <- data.frame(p = seq(.3, .8, by=.1), 
                          Prior = c(0.125, 0.125, 0.250, 
                                    0.250, 0.125, 0.125))
ggplot(data=bayes_table, aes(x=p, y=Prior)) +
  geom_bar(stat="identity", fill=crcblue, width = 0.06)
```

## Step 2: Data/likelihood of proportion $p$
- Now the Tokyo Express owner gives a survey to 20 customers.
- Out of the 20 responses, 12 say that their favorite day for eating out for dinner is Friday.
\pause
- Quantity of interest: $p$, the proportion of customers prefer eating out for dinner on Friday.
\pause
- The data/likelihood is a function of the quantity of interest.
- What would be the function of 12 out of 20 preferring Friday, in terms of the proportion $p$?

## The Binomial distribution
- A Binomial experiment:
    1. One is repeating the same basic task or trial many times -- let the number of trials be denoted by $n$.
    2. On each trial, there are two possible outcomes that are  called ``success" or ``failure".
    3. The probability of a success, denoted by $p$, is the same for each trial.
    4. The results of outcomes from different trials are independent.
- Do you think the survey is a Binomial experiment?
\pause
- The probability of $y$ successes in a Binomial experiment is given by
\begin{eqnarray}
P(Y=y) = {n \choose y} p^y (1 - p)^{n - y}, y = 0, \cdots, n,
\end{eqnarray}
where $n$ is the number of trials and $p$ is the success probability.

## The likelihood function
- The probability of $y$ successes in a binomial experiment is given by
\begin{eqnarray}
P(Y=y) = {n \choose y} p^y (1 - p)^{n - y}, y = 0, \cdots, n,
\end{eqnarray}
where $n$ is the number of trials and $p$ is the success probability.
\pause
- The likelihood is the chance of 12 successes in 20 trials viewed as a function of the probability of success is $p$:
\begin{eqnarray}
\textrm{Likelihood} = L(p) = {20 \choose 12} p ^ {12} (1 - p) ^ 8.
\label{eq:dining:likelihood}
\end{eqnarray}
    - $L$ is a function of $p$.
    - $n$ is fixed and known.
    - $Y$ is the random variable.
    - $p$ is the quantity of interest, also the unknown parameter in the Binomial distribution.

## Use R/RStudio to compute the likelihood function
\begin{eqnarray}
\textrm{Likelihood} = L(p) = {20 \choose 12} p ^ {12} (1 - p) ^ 8
\label{eq:dining:likelihood}
\end{eqnarray}

- Need: sample size $n$ (20), number of successes $k$ (12), and possible values of proportion $p$ (\{0.3, 0.4, 0.5, 0.6, 0.7, 0.8\}).
- Do not need: the assigned probabilities (0.125, 0.125, 0.250, 0.250, 0.125, 0.125) in the prior distribution $\pi_{owner}(p)$.

## Use R/RStudio to compute the likelihood function 
```{r, size = "footnotesize"}
bayes_table$Likelihood <- dbinom(12, size=20, 
                                 prob=bayes_table$p)
bayes_table
```

## Step 3: Posterior distribution
- Notations:
    - $\pi(p)$ the prior distribution of $p$.
    - $L(p)$ is the likelihood function.
    - $\pi(p \mid y)$ the posterior distribution of $p$ after observing the number of successes $y$.
\pause
- The Bayes' rule for a discrete parameter has the form
\begin{eqnarray}
\pi(p_i \mid y)  = \frac{\pi(p_i) \times L(p_i)} {\sum_j \pi(p_j) \times L(p_j)},
\label{eq:Discrete:bayesrule}
\end{eqnarray}
    - $\pi(p_i)$ the prior probability of $p = p_i$.
    - $L(p_i)$ the likelihood function evaluated at $p = p_i$.
    - $\pi(p_i \mid y)$ the posterior probability of $p = p_i$ given the number of successes $y$.
    - the denominator gives the marginal distribution of the observation $y$ (by the $\color{VassarRed}{\text{Law of Total Probability}}$).
    
## Use R/RStudio to compute and plot the posterior
```{r, size = "footnotesize"}
bayesian_crank(bayes_table) -> bayes_table
bayes_table
```

- Inference question: What is the $\color{VassarRed}{\text{posterior probability}}$ that over half of the customers prefer eating out on Friday for dinner?
\pause
\begin{eqnarray}
 Prob(p > 0.5) = 0.463 + 0.147 + 0.029 = 0.639.
\end{eqnarray}

## Use R/RStudio to compute and plot the posterior
```{r, size = "footnotesize"}
bayesian_crank(bayes_table) -> bayes_table
sum(bayes_table$Posterior[bayes_table$p > 0.5])
```
\pause

- Exercise: What is the $\color{VassarRed}{\text{posterior probability}}$ that less than 40\% of the customers prefer eating out on Friday for dinner? 

## Using R/RStudio to compute and plot the posterior
```{r fig.height = 2.7, fig.width = 2.7, size = "footnotesize", fig.align = "center"}
prior_post_plot(bayes_table, Color = crcblue) +
  theme(text=element_text(size=10))
```

# Continuous priors - the Beta distribution

## Step 1: Prior distribution
- $\color{VassarRed}{\text{Before giving out the survey}}$, we need to specify a prior distribution for unknown parameter $p$.
- Previously, $p$ can take 6 possible values
\begin{eqnarray}
p = \{0.3, 0.4, 0.5, 0.6, 0.7, 0.8\}.
\label{eq:dining:p}
\end{eqnarray}
- And a discrete prior distribution:
\begin{eqnarray}
\pi_{owner}(p)= (0.125, 0.125, 0.250, 0.250, 0.125, 0.125).
\label{eq:dining:expertprior}
\end{eqnarray}

## Step 1: Prior distribution
```{r, size = "footnotesize"}
bayes_table
```

- Anything unsatisfactory?

## Move to continuous priors
- A limitation of specifying a discrete prior for $p$
    - If a plausible value is not specified in the prior distribution (e.g. $p = 0.2$), it will be assigned a 0 probability in the posterior distribution.
\pause
- Ideally, we want a distribution that allows $p$ to be any value in [0, 1].
- The continuous Uniform distribution:
    - Any value of $p$ is equally likely.
    - The probability density function of the continuous Uniform on the interval $[a, b]$ is
\begin{eqnarray}
\pi(p) = 
\begin{cases}
  \frac{1}{b - a} & \text{for }a \le p \le b,\\    
  0     		& \text{for }p < a \,\, \text{or } p > b.
\end{cases}
\label{eq:Binomial:Continuous:Uniform}
\end{eqnarray}
    - $p \sim \textrm{Uniform}(0,1)$: a very special case of $p$.
- The Beta distribution!

## The Beta distribution
- Notation: $\textrm{Beta}(a, b)$.
- For a random variable falling between 0 and 1, suitable for proportion $p$.
- Beta distribution has two shape parameters $a$ and $b$.
- Probability density function (pdf) is:
\begin{eqnarray}
\pi(p) = \frac{1}{B(a, b)} p^{a - 1} (1 - p)^{b - 1}, \, \, 0 \le p \le 1.
\end{eqnarray}
    - $B(a, b)$ is the Beta function $B(a, b) = \frac{\Gamma(a)\Gamma(b)}{\Gamma(a+b)}$.
    - $\Gamma$ is the Gamma function.
    \pause
    - Continuous Uniform on [0, 1] is a special case of Beta with $a=b=1$: $\textrm{Uniform}(0, 1) = \textrm{Beta}(1, 1)$

## Examples of Beta curves
```{r, echo = FALSE, fig.align = "center"}
betapars <- matrix(c(0.5, 0.5,
                     0.5, 1,
                     0.5, 2,
                     1, 0.5,
                     1, 1,
                     1, 2,
                     4, 0.5,
                     4, 1,
                     4, 2),
                     9, 2, byrow = TRUE)
p <- seq(.001, .999, length.out = 100)
BETA <- NULL
for (j in 1:9){
  df <- data.frame(p = p, Density = dbeta(p,
                    betapars[j, 1], betapars[j, 2]))
  df$Type <-  paste("Beta(", betapars[j, 1],
                    ",", betapars[j, 2], ")",
                    sep = "")
  BETA <- rbind(BETA, df)
}
ggplot(BETA, aes(p, Density)) +
  geom_line(color = crcblue, size = 1.5) +
  facet_wrap(~ Type, scale = "free") +
  increasefont() +
  theme(axis.text=element_blank())
```

## Choose a Beta curve to represent prior opinion
- Prior opinion: values of $p$ and associated probabilities.
- Difficult to guess values of $a$ and $b$ in $\textrm{Beta}(a,b)$.
- Solution: specifying a Beta prior by specification of quantiles of the distribution.
\pause

    - Quantiles are about rank order of values.
    - e.g. Middle quantile/50-th percentile: median.
    - \texttt{beta\_quantile()} function in R: inputs a probability measure $p$ and outputs the value of $x$ such that $Prob(X <=x) = p$. (e.g. $x = 0.408$ when $p = 0.5$ for $\textrm{Beta}(7, 10)$)

## Choose a Beta curve to represent prior opinion
```{r fig.height = 2.8, fig.width = 2.8, size = "footnotesize", fig.align = "center"}
beta_quantile(0.5, c(7, 10), Color = crcblue) +
  theme(text=element_text(size=8))
```

## Use \texttt{beta.select()} to choose a Beta curve
- The \texttt{beta.select()} function needs us to 
    - first think about specifying two quantiles
    - then it finds the Beta curve that matches with these quantiles
    \pause

- Example:
    - Suppose we believe that $p_{50} = 0.55$ (50-th quantile)
    - Suppose we also believe that $p_{90} = 0.8$ (90-th quantile)
    - Input these two sets of values into \texttt{beta.select()}
    
```{r, size = "footnotesize"}
beta.select(list(x = 0.55, p = 0.5),
            list(x = 0.80, p = 0.9))
```

\pause

- Exercise 1: Verify that $\textrm{Beta}(3.06, 2.56)$ has $p_{50} = 0.55$ and $p_{90} = 0.8$. (Hint: use the \texttt{beta\_quantile()} function)
- Exercise 2: Come up with your own Beta prior distribution, and share it with your neighbors.

# Updating the Beta prior

## Step 2: Data/likelihood of proportion $p$
- Recall that
    - The Tokyo Express owner gives a survey to 20 customers, and 12 respond that their favorite day is Friday.
    - The data/likelihood is a function of the quantity of interest, $p$.
    - It is a Binomial experiment, and 
\begin{eqnarray}
P(Y=y) = {n \choose y} p^y (1 - p)^{n - y}, y = 0, \cdots, n,
\end{eqnarray}
where $n$ is the number of trials and $p$ is the success probability.
\pause
- Exercise: Write out the likelihood function for $n = 20$ and $y = 12$.
\pause
- Solution:
\begin{eqnarray}
\textrm{Likelihood} = L(p) = {20 \choose 12} p ^ {12} (1 - p) ^ 8.
\end{eqnarray}

## Bayes' rule for continuous priors
- The likelihood function is the same, regardless of the prior distribution.
- Recall: the Bayes' rule for a discrete parameter has the form
\begin{eqnarray}
\pi(p_i \mid y)  = \frac{\pi(p_i) \times L(p_i)} {\sum_j \pi(p_j) \times L(p_j)}
\end{eqnarray}
- What about for continuous $p$? Unfortunately we can not list each value of $p$ anymore.
\pause
- Solution: With continuous $p$, the denominator changes from summation $\Sigma$ to integration $\int$.
- Since $\int \pi(p) \times L(p) dp = f(y)$ is fixed, we can write the Bayes' rule for continuous $p$ in proportional sign:
\begin{eqnarray}
\pi(p \mid y) \propto  \pi(p) \times L(p) .
\end{eqnarray}

## Step 3: Derive the posterior
\begin{eqnarray}
\pi(p \mid y) \propto  \pi(p) \times L(p)
\end{eqnarray}

- For prior $\pi(p)$, we have $p \sim \textrm{Beta}(3.06, 2.56)$.
- For data/likelihood $L(p)$, we have $Y \sim \textrm{Binomial}(20,p)$.
- We need to derive $\pi(p \mid y)$.
\pause

- Exercise: Derive $\pi(p \mid y)$ with the setup below.
    - The prior distribution:
      \begin{eqnarray}
      \pi(p) = \frac{1}{B(3.06, 2.56)}p^{3.06-1}(1-p)^{2.56-1}. \nonumber
      \end{eqnarray}
    - The likelihood:
      \begin{eqnarray}
      f(Y =12 \mid p) = L(p) = {20 \choose 12}p^{12}(1-p)^{8}.\nonumber
      \end{eqnarray}
    - The posterior?
    
## Step 3: Derive the posterior cont'd
- The posterior:
  \begin{eqnarray}
  \pi(p \mid Y = 12) &\propto& \pi(p) \times f(Y = 12 \mid p) \nonumber \\ \pause
  &=&  \frac{1}{B(3.06, 2.56)}p^{3.06-1}(1-p)^{2.56-1} \times \nonumber \\ 
  && {20 \choose 12}p^{12}(1-p)^{8} \nonumber \\ \pause
  \texttt{[drop the constants]} &\propto& p^{12}(1-p)^{8}p^{3.06-1}(1-p)^{2.56-1} \nonumber \\
  \texttt{[combine the powers]} &=& p^{15.06-1}(1-p)^{10.56-1}.
  \end{eqnarray}
- That is, 
  \begin{eqnarray}
  \pi(p \mid Y = 12) \propto p^{15.06-1}(1-p)^{10.56-1},\nonumber
  \end{eqnarray}
  which means \pause
  \begin{eqnarray}
  p \mid Y = 12 \sim \textrm{Beta}(15.06, 10.56).\nonumber
\end{eqnarray}

## From Beta prior to Beta posterior: conjugate priors
- The prior distribution:
  \begin{eqnarray}
  p \sim \textrm{Beta}(a, b)\nonumber
  \end{eqnarray}
- The sampling density:
  \begin{eqnarray}
  Y \sim \textrm{Binomial}(n, p)\nonumber
  \end{eqnarray}
- The posterior distribution:
  \begin{eqnarray}
  p \mid Y = y \sim \textrm{Beta}(a + y, b + n - y)\nonumber
  \end{eqnarray}
- Conjugate priors: from Beta prior to Beta posterior.
\pause
\begin{table}[htb]
%\caption{\label{table:betaupdate} Updating the beta prior.}
\begin{center}
\begin{tabular}{|c|c|c|} \hline
Source & Successes & Failures \\ \hline
Prior & $a$ & $b$ \\
Data/Likelihood & $y$  & $n-y$ \\
Posterior & $a + y$ & $b + n - y$ \\ \hline
\end{tabular}
\end{center}
\label{default}
\end{table}

## Use R/RStudio to compute and plot the posterior
\begin{eqnarray}
p \mid Y = y \sim \textrm{Beta}(a + y, b + n - y)\nonumber
\end{eqnarray}


```{r, size = "footnotesize"}
ab <- c(3.06, 2.56)
yny <- c(12, 8)
(ab_new <- ab + yny)
```

## Use R/RStudio to compute and plot the posterior cont'd

\columnsbegin
\column{.6\textwidth}
```{r echo = FALSE, fig.height=5, fig.width=5}
ggplot(data.frame(x=c(0, 1)), aes(x)) +
  stat_function(fun=dbeta, geom="line",
                aes(linetype="solid"),
                size=1.5, color = crcblue,
                args=list(shape1=ab[1],
                          shape2=ab[2])) +
  stat_function(fun=dbeta, geom="line",
                aes(linetype="dashed"),
                size=1.5, color = crcblue,
                args=list(shape1=ab_new[1],
                          shape2=ab_new[2])) +
  xlab("p") + ylab("Density") +
  theme(legend.position = "none") +
  annotate(geom = "text", x = .15, y = 1.2,
           size = 8, label = "Prior") +
  annotate(geom = "text", x = .85, y = 3,
           size = 8, label = "Posterior") +
  increasefont()
```
\column{.3\textwidth}
\pause

- Exercise 1: Compare prior mean 0.544 and posterior mean 0.588. (Recall that sample mean is 0.6.)
- Exercise 2: Compare the spreads of the two curves.
    
\columnsend


# Bayesian inference with continuous priors

## Bayesian hypothesis testing
- Suppose one of the Tokyo Express's workers claims that at least 75\% of the customers prefer Friday. Is this a reasonable claim?
\pause
- From a Bayesian viewpoint, 
    - Find the posterior probability that $p >= 0.75$.
    - Make a decision based on the value of the posterior probability.
    - If the probability is small, we can reject this claim.

## Bayesian hypothesis testing cont'd
\columnsbegin
\column{.6\textwidth}
```{r, size = "footnotesize"}
beta_area(lo = 0.75, hi = 1.0, shape_par = c(15.06, 10.56), 
          Color = crcblue) +
  theme(text=element_text(size=18))
```

\column{.3\textwidth}
\pause

- Posterior probability is only 4\%, reject the claim.
- Exercise: What about a claim ``at most 30\% of the customers prefer Friday"?
    
\columnsend

## Bayesian hypothesis testing cont'd
When the posterior distribution is known...
\pause

- Exact solution: use the \texttt{beta\_area()} function in the \texttt{ProbBayes} package and/or the \texttt{pbeta()} R function
    
```{r, eval = FALSE, size = "footnotesize"}
beta_area(lo = 0.75, hi = 1.0, shape_par = c(15.06, 10.56))
```
    
```{r, size = "footnotesize"}
pbeta(1, 15.06, 10.56) - pbeta(0.75, 15.06, 10.56)
```

\pause

- Approximation through Monte Carlo simulation using the \texttt{rbeta()} R function

```{r, size = "footnotesize"}
S <- 1000
BetaSamples <- rbeta(S, 15.06, 10.56)
sum(BetaSamples >= 0.75)/S
```

## Bayesian credible intervals
- Consider an interval that we are confident contains $p$.
- Such an interval provides an uncertainty estimate for our parameter $p$.
- A 90\% Bayesian credible interval is an interval contains 90\% of the posterior probability.

## Bayesian credible intervals cont'd
\columnsbegin
\column{.6\textwidth}
```{r, size = "footnotesize", fig.align = "center"}
beta_interval(0.9, c(15.06, 10.56), Color = crcblue) +
  theme(text=element_text(size=18))
```
\column{.3\textwidth}
\pause

- The middle 90\%.
- Interpretation: different from a confidence interval.
- Exercise: Construct a middle 98\% credible interval for $p$.
    
\columnsend

## Bayesian credible intervals cont'd
When the posterior distribution is known...
\pause

- Exact solution: use the \texttt{beta\_interval()} function in the \texttt{ProbBayes} package ($\color{VassarRed}{\text{only the middle 90\%}}$) and or the \texttt{qbeta()} R function ($\color{VassarRed}{\text{not necessarily the middle 90\%}}$)
    
```{r, eval = FALSE, size = "footnotesize"}
beta_interval(0.9, c(15.06, 10.56), Color = crcblue)
```

```{r, size = "footnotesize"}
c(qbeta(0.05, 15.06, 10.56), qbeta(0.95, 15.06, 10.56))
```

\pause

- Approximation through Monte Carlo simulation using the \texttt{rbeta()} R function ($\color{VassarRed}{\text{not necessarily the middle 90\%}}$)
  
```{r, size = "footnotesize"}
S <- 1000; BetaSamples <- rbeta(S, 15.06, 10.56)
quantile(BetaSamples, c(0.05, 0.95))
```
    
    
## Bayesian prediction
- Suppose the Tokyo Express owner gives out another survey to $m$ customers, how many would prefer Friday?
- The predictive distribution: $\tilde{Y} \mid Y = y$.
\pause

    - The exact prediction:
    \begin{eqnarray*}
    \tilde{Y} \mid Y = y \sim \textrm{Beta-Binomial}(m, a + y, b + n - y).
    \end{eqnarray*}
    
    - Prediction through simulation ($\color{VassarRed}{\text{our focus}}$):
    \begin{eqnarray*}
    \text{sample}\,\, p \sim {\rm{Beta}}(a + y, b + n - y) &\rightarrow& \text{sample}\,\,         \tilde{y} \sim {\rm{Binomial}}(m, p)\\
    \end{eqnarray*}

## Use R/RStudio to make Bayesian predictions
```{r, size = "footnotesize"}
S <- 1000
a <- 3.06; b <- 2.56
n <- 20; y <- 12
m <- 20
pred_p_sim <- rbeta(S, a + y, b + n - y)
pred_y_sim <- rbinom(S, m, pred_p_sim)
sum(pred_y_sim >=5 & pred_y_sim <= 15)/S
```

## Use R/RStudio to make Bayesian predictions

```{r, echo = FALSE, warning = FALSE}
a <- 3.06; b <- 2.56
n <- 20; y <- 12
prob <- pbetap(c(a + y, b + n - y), 20, 0:20)
T1 <- data.frame(Y = 0:10,
                 Probability = round(prob[1:11], 3))
T2 <- data.frame(Y = 11:20,
                 Probability = round(prob[12:21], 3))
T2 <- rbind(T2, data.frame(Y = 21, Probability = 999))

set.seed(123)
S <- 1000
pred_p_sim <- rbeta(S, a + y, a + b + n - y)
pred_y_sim <- rbinom(S, n, pred_p_sim)

data.frame(Y = pred_y_sim) %>%
  group_by(Y) %>% summarize(N = n()) %>%
  mutate(Probability = N / sum(N),
         Type = "Simulated")  %>%
  select(Type, Y, Probability) -> S1

S2 <- data.frame(Type = "Exact",
                 Y = 0:20,
                 Probability = prob)

S <- rbind(S1, S2)
ggplot(S, aes(Y, Probability)) +
  geom_segment(aes(xend = Y, yend = 0),
               size = 3,
               lineend = "butt",
               color = crcblue) +
  facet_wrap(~ Type, ncol=1) +
  theme(text=element_text(size=18))
```

## Bayesian predictive checking
- Bayesian prediction is $\color{VassarRed}{\text{posterior predictive}}$
\small
\begin{eqnarray*}
\text{sample}\,\, p \sim {\rm{Beta}}(a + y, b + n - y) &\rightarrow& \text{sample}\,\, \tilde{y} \sim {\rm{Binomial}}(m, p)\\
\end{eqnarray*}
- We can perform $\color{VassarRed}{\text{posterior predictive checking}}$ through simulation
\begin{eqnarray*}
\text{sample}\,\, p^{(1)} \sim {\rm{Beta}}(a + y, b + n - y) &\rightarrow& \text{sample}\,\, \tilde{y}^{(1)} \sim {\rm{Binomial}}({\color{red}n}, p^{(1)})\\
\text{sample}\,\, p^{(2)} \sim {\rm{Beta}}(a + y, b + n - y) &\rightarrow& \text{sample}\,\, \tilde{y}^{(2)} \sim {\rm{Binomial}}({\color{red}n}, p^{(2)})\\
&\vdots& \\
\text{sample}\,\, p^{(S)} \sim {\rm{Beta}}(a + y, b + n - y) &\rightarrow& \text{sample}\,\, \tilde{y}^{(S)} \sim {\rm{Binomial}}({\color{red}n}, p^{(S)})\\
\end{eqnarray*}

\normalsize
The sample  $\{\tilde{y}^{(1)}, ..., \tilde{y}^{(S)}\}$ is an approximation to the posterior predictive distribution that can be used for model checking. 

## Use R/Rstudio to perform posterior predictive checking
```{r, size = "footnotesize"}
S <- 1000
a <- 3.06; b <- 2.56
n <- 20; y <- 12
newy = as.data.frame(rep(NA, S))
names(newy) = c("y")

set.seed(123)
for (s in 1:S){
  pred_p_sim <- rbeta(1, a + y, b + n - y)
  pred_y_sim <- rbinom(1, n, pred_p_sim)
  newy[s,] = pred_y_sim 
}

```

## Use R/Rstudio to perform posterior predictive checking cont'd
\columnsbegin
\column{.6\textwidth}
```{r, size = "footnotesize", fig.height = 2, fig.width = 2, fig.align = "center"}
ggplot(data=newy, aes(newy$y)) + 
  geom_histogram(breaks=seq(0, 20, by=0.5), fill = crcblue) +
  annotate("point", x = 12, y = 0, colour = "red", size = 5) +
  xlab("y") + theme(text=element_text(size=10))
```

\column{.4\textwidth}
\pause

- The observed $y = 12$ is plotted as a red dot. The observed value of $y$ is consistent with simulations of replicated data from this predictive distribution.
  
\columnsend

## Use R/Rstudio to perform posterior predictive checking cont'd
- More formally, one can calculate the following probabilities:

\begin{equation}
Prob(y > \tilde{y} \mid y), \,\,\, \text{or}\,\,\, 1 - Prob(y > \tilde{y} \mid y).
\end{equation}

- If either probability is small, it suggests the model does not describe $y$ very well.
\pause

```{r, size = "footnotesize"}
sum(newy > y)/S
1 - sum(newy > y)/S
```

- Since 0.407 and 0.593 are not small, it suggests the model describe $y$ well. The inference passes the posterior predictive checking. 


# Recap

## Recap
- Bayesian inference procedure:
    - Step 1: express an opinion about the location of the proportion $p$ before sampling (prior).
    - Step 2: take the sample and record the observed proportion (data/likelihood).
    - Step 3: use Bayes' rule to sharpen and update the previous opinion about $p$ given the information from the sample (posterior).

\pause

- For Binomial data/likelihood, the Beta distributions are conjugate priors.
    - The prior distribution: $p \sim \textrm{Beta}(a, b)$.
    - The sampling density: $Y \sim \textrm{Binomial}(n, p)$.
    - The posterior distribution: $p \mid Y = y \sim \textrm{Beta}(a + y, b + n - y)$.
   
\pause

- Bayesian inferences (exact vs simulated)
    - Bayesian hypothesis testing \& Bayesian credible intervals
    - Bayesian predictions
    - Bayesian posterior predictive checking