---
title: "Bayesian linear regression"
author: "Jingchen (Monika) Hu"
date: "MATH 347 Bayesian Statistics"
output:
  pdf_document: default
  html_document:
    number_sections: yes
institute: Vassar College
fontsize: 11pt
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
require(gridExtra)
require(ProbBayes)
require(tidyverse)
require(runjags)
require(coda)
crcblue <- "#2905a1"
knitr::opts_chunk$set(echo = TRUE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

# Installing the necessary packages

```{r, eval = FALSE}
install.packages("devtools")
require(devtools)
devtools::install_github("bayesball/ProbBayes")

require(ggplot2)
require(gridExtra)
require(ProbBayes)
require(tidyverse)
crcblue <- "#2905a1"
```



# Introduction: Adding a continuous predictor variable

## The simple linear regression model

```{r fig.align = "center"}
CEData <- read.csv("CEsample.csv", header = T, sep = ",")
g1 <- ggplot(CEData, aes(x = log_TotalIncome, y = log_TotalExp)) +
  geom_point(size=1) + 
  labs(x = "log(Income)", y = "log(Expenditure)") +
  theme_grey(base_size = 10, base_family = "") 
g1
```


# The CE sample


# A simple linear regression for the CE sample

# MCMC simulation by JAGS for the SLR model

## JAGS script for the SLR model

```{r message = FALSE}
modelString <-"
model {
## sampling
for (i in 1:N){
y[i] ~ dnorm(beta0 + beta1*x[i], invsigma2)
}

## priors
beta0 ~ dnorm(mu0, g0)
beta1 ~ dnorm(mu1, g1)
invsigma2 ~ dgamma(a, b)
sigma <- sqrt(pow(invsigma2, -1))
}
"
```

- Pass the data and hyperparameter values to JAGS:

```{r message = FALSE}
y <- as.vector(CEData$log_TotalExp)
x <- as.vector(CEData$log_TotalIncome)
N <- length(y)
the_data <- list("y" = y, "x" = x, "N" = N,
                 "mu0" = 0, "g0" = 0.0001,
                 "mu1" = 0, "g1" = 0.0001,
                 "a" = 1, "b" = 1)

initsfunction <- function(chain){
  .RNG.seed <- c(1,2)[chain]
  .RNG.name <- c("base::Super-Duper",
                 "base::Wichmann-Hill")[chain]
  return(list(.RNG.seed=.RNG.seed,
              .RNG.name=.RNG.name))
}
```


- Run the JAGS code for this model:

```{r message = FALSE, warning = FALSE}
posterior <- run.jags(modelString,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("beta0", "beta1", "sigma"),
                      adapt = 1000,
                      burnin = 5000,
                      sample = 5000,
                      thin = 1,
                      inits = initsfunction)
```


## JAGS output for the SLR model

- Obtain posterior summaries of all parameters:

```{r message = FALSE, warning = FALSE}
summary(posterior) 
```


```{r fig.align = "center", message = FALSE}
plot(posterior, vars = "beta0")
```


```{r fig.align = "center", message = FALSE}
plot(posterior, vars = "beta1")
```

```{r fig.align = "center", message = FALSE}
plot(posterior, vars = "sigma")
```


## New JAGS script for the SLR model

Setting \texttt{thin = 50}, to get rid of the stickiness in $\beta_0$ and $\beta_1$.

```{r message = FALSE,warning = FALSE}
posterior_new <- run.jags(modelString,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("beta0", "beta1", "sigma"),
                      adapt = 1000,
                      burnin = 5000,
                      sample = 5000,
                      thin = 50,
                      inits = initsfunction)
```


## New JAGS output for the SLR model

- Obtain posterior summaries of all parameters:


```{r message = FALSE, warning = FALSE}
summary(posterior_new) 
```


```{r fig.align = "center", message = FALSE}
plot(posterior_new, vars = "beta0")
```


```{r fig.align = "center", message = FALSE}
plot(posterior_new, vars = "beta1")
```


```{r fig.align = "center", message = FALSE}
plot(posterior_new, vars = "sigma")
```


# Bayesian inferences with SLR

## Simulate fits from the regression model


```{r message = FALSE, warning = FALSE}
post <- as.mcmc(posterior_new)
post_means <- apply(post, 2, mean)
post <- as.data.frame(post)
```

```{r fig.align = "center", message = FALSE}
ggplot(CEData, aes(log_TotalIncome, log_TotalExp)) +
  geom_point(size=1) +
  geom_abline(data=post[1:10, ],
              aes(intercept=beta0, slope=beta1), alpha = 0.5) +
  geom_abline(intercept = post_means[1],
              slope = post_means[2], size = 1) +
  ylab("log(Expenditure)") + xlab("log(Income)") +
  theme_grey(base_size = 10, base_family = "")
```


## Learning about the expected response


```{r message = FALSE, warning = FALSE}
post <- as.data.frame(post)
one_expected <- function(x){
  lp <- post[ , "beta0"] +  x * post[ , "beta1"]
  data.frame(Value = paste("log(Income) =", x),
             Expected_logExp = lp)
}

df <- map_df(c(1, 5, 7, 9), one_expected)
```


```{r fig.align = "center", message = FALSE}
require(ggridges)
ggplot(df, aes(x = Expected_logExp, y = Value)) +
  geom_density_ridges() +
  theme_grey(base_size = 8, base_family = "")
```

```{r message = FALSE, warning = FALSE}
df_new <- map_df(c(10, 11, 12, 13), one_expected)
ggplot(df_new, aes(x = Expected_logExp, y = Value)) +
  geom_density_ridges() +
  theme_grey(base_size = 8, base_family = "")
```

```{r message = FALSE, warning = FALSE}
df %>% group_by(Value) %>%
  summarize(P05 = quantile(Expected_logExp, 0.05),
            P50 = median(Expected_logExp),
            P95 = quantile(Expected_logExp, 0.95))
```


## Prediction of future responses

```{r message = FALSE, warning = FALSE}
one_predicted <- function(x){
  lp <- post[ , "beta0"] +  x * post[ , "beta1"]
  y <- rnorm(5000, lp, post[, "sigma"])
  data.frame(Value = paste("log(Income) =", x),
             Predicted_logExp = y)
}
df <- map_df(c(1, 5, 7, 9), one_predicted)
```


```{r fig.align = "center", message = FALSE}
require(ggridges)
ggplot(df, aes(x = Predicted_logExp, y = Value)) +
  geom_density_ridges() +
  theme_grey(base_size = 9, base_family = "")
```



```{r message = FALSE, warning = FALSE}
df %>% group_by(Value) %>%
  summarize(P05 = quantile(Predicted_logExp, 0.05),
            P50 = median(Predicted_logExp),
            P95 = quantile(Predicted_logExp, 0.95))
```


# More on priors

## Subjective prior: standardization

```{r message = FALSE, warning = FALSE}
CEData$log_TotalExpSTD <- scale(CEData$log_TotalExp)
CEData$log_TotalIncomeSTD <- scale(CEData$log_TotalIncome)
```


```{r fig.align = "center", message = FALSE}
g2 = ggplot(CEData, aes(x = log_TotalIncomeSTD, y = log_TotalExpSTD)) +
  geom_point(size=1) + 
  xlab("log(Income) STD") + ylab("log(Expenditure) STD") +
  theme_grey(base_size = 10, base_family = "") 
grid.arrange(g1, g2, ncol=2)
```

## Subjective prior: JAGS script for the standardized SLR model

```{r message = FALSE}
modelString <-"
model {
## sampling
for (i in 1:N){
y[i] ~ dnorm(beta0 + beta1*x[i], invsigma2)
}

## priors
beta0 ~ dnorm(mu0, g0)
beta1 ~ dnorm(mu1, g1)
invsigma2 ~ dgamma(a, b)
sigma <- sqrt(pow(invsigma2, -1))
}
"
```

- Pass the data and hyperparameter values to JAGS:

```{r message = FALSE}
y <- as.vector(CEData$log_TotalExpSTD)
x <- as.vector(CEData$log_TotalIncomeSTD)
N <- length(y)
the_data <- list("y" = y, "x" = x, "N" = N,
                 "mu0" = 0, "g0" = 1,
                 "mu1" = 0.7, "g1" = 44.4,
                 "a" = 1, "b" = 1)

initsfunction <- function(chain){
  .RNG.seed <- c(1,2)[chain]
  .RNG.name <- c("base::Super-Duper",
                 "base::Wichmann-Hill")[chain]
  return(list(.RNG.seed=.RNG.seed,
              .RNG.name=.RNG.name))
}
```



- Run the JAGS code for this model:

```{r message = FALSE, warning = FALSE}
posterior_sub <- run.jags(modelString,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("beta0", "beta1", "sigma"),
                      adapt = 1000,
                      burnin = 5000,
                      sample = 5000,
                      thin = 1,
                      inits = initsfunction)
```

## Subjective prior: JAGS output for the SLR model

- Obtain posterior summaries of all parameters:

\vspace{3mm}

```{r message = FALSE, warning = FALSE}
summary(posterior_sub) 
```


```{r fig.align = "center", message = FALSE}
plot(posterior_sub, vars = "beta1")
```

## Conditional means prior: JAGS script

```{r message = FALSE}
modelString <-"
model {
## sampling
for (i in 1:N){
y[i] ~ dnorm(beta0 + beta1*x[i], invsigma2)
}

## priors
beta1 <- (mu2 - mu1)/(x2 - x1)
beta0 <- mu1 - x1*(mu2 - mu1)/(x2 - x1)
mu1 ~ dnorm(m1, g1)
mu2 ~ dnorm(m2, g2)
invsigma2 ~ dgamma(a, b)
sigma <- sqrt(pow(invsigma2, -1))
}
"
```



# A multiple linear regression, and MCMC simulation by JAGS

## JAGS script for the MLR model

```{r message = FALSE}
CEData$log_TotalExpSTD <- scale(CEData$log_TotalExp)
CEData$log_TotalIncomeSTD <- scale(CEData$log_TotalIncome)

library(fastDummies)
## create indictor variable for Rural
CEData$Rural = fastDummies::dummy_cols(CEData$UrbanRural)[,names(fastDummies::dummy_cols(CEData$UrbanRural))
 == ".data_2"]
```

```{r message = FALSE}
## create indicator variables for Black (2), Native American (3), 
## Asian (4), Pacific Islander (5), and Multi-race (6)
CEData$Race_Black = fastDummies::dummy_cols(CEData$Race)[,names(fastDummies::dummy_cols(CEData$Race)) == ".data_2"]
CEData$Race_NA = fastDummies::dummy_cols(CEData$Race)[,names(fastDummies::dummy_cols(CEData$Race)) == ".data_3"]
CEData$Race_Asian = fastDummies::dummy_cols(CEData$Race)[,names(fastDummies::dummy_cols(CEData$Race)) == ".data_4"]
CEData$Race_PI = fastDummies::dummy_cols(CEData$Race)[,names(fastDummies::dummy_cols(CEData$Race)) == ".data_5"]
CEData$Race_M = fastDummies::dummy_cols(CEData$Race)[,names(fastDummies::dummy_cols(CEData$Race)) == ".data_6"]
```

```{r message = FALSE}
modelString <-"
model {
## sampling
for (i in 1:N){
y[i] ~ dnorm(beta0 + beta1*x_income[i] + beta2*x_rural[i] +
beta3*x_race_B[i] + beta4*x_race_N[i] +
beta5*x_race_A[i] + beta6*x_race_P[i] +
beta7*x_race_M[i], invsigma2)
}
## priors
beta0 ~ dnorm(mu0, g0)
beta1 ~ dnorm(mu1, g1)
beta2 ~ dnorm(mu2, g2)
beta3 ~ dnorm(mu3, g3)
beta4 ~ dnorm(mu4, g4)
beta5 ~ dnorm(mu5, g5)
beta6 ~ dnorm(mu6, g6)
beta7 ~ dnorm(mu7, g7)
invsigma2 ~ dgamma(a, b)
sigma <- sqrt(pow(invsigma2, -1))
}
"
```

- Pass the data and hyperparameter values to JAGS:

```{r message = FALSE}
y = as.vector(CEData$log_TotalExpSTD)
x_income = as.vector(CEData$log_TotalIncomeSTD)
x_rural = as.vector(CEData$Rural)
x_race_B = as.vector(CEData$Race_Black)
x_race_N = as.vector(CEData$Race_NA)
x_race_A = as.vector(CEData$Race_Asian)
x_race_P = as.vector(CEData$Race_PI)
x_race_M = as.vector(CEData$Race_M)
N = length(y)  # Compute the number of observations
```

- Pass the data and hyperparameter values to JAGS:

```{r message = FALSE}
the_data <- list("y" = y, "x_income" = x_income,
                 "x_rural" = x_rural, "x_race_B" = x_race_B,
                 "x_race_N" = x_race_N, "x_race_A" = x_race_A,
                 "x_race_P" = x_race_P, "x_race_M" = x_race_M,
                 "N" = N,
                 "mu0" = 0, "g0" = 1, "mu1" = 0, "g1" = 1,
                 "mu2" = 0, "g2" = 1, "mu3" = 0, "g3" = 1,
                 "mu4" = 0, "g4" = 1, "mu5" = 0, "g5" = 1,
                 "mu6" = 0, "g6" = 1, "mu7" = 0, "g7" = 1,
                 "a" = 1, "b" = 1)
```

- Pass the data and hyperparameter values to JAGS:

```{r message = FALSE, size = "footnotesize"}
initsfunction <- function(chain){
  .RNG.seed <- c(1,2)[chain]
  .RNG.name <- c("base::Super-Duper",
                 "base::Wichmann-Hill")[chain]
  return(list(.RNG.seed=.RNG.seed,
              .RNG.name=.RNG.name))
}
```

- Run the JAGS code for this model:

```{r message = FALSE}
posterior_MLR <- run.jags(modelString,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("beta0", "beta1", "beta2",
                                  "beta3", "beta4", "beta5",
                                  "beta6", "beta7", "sigma"),
                      adapt = 1000,
                      burnin = 5000,
                      sample = 5000,
                      thin = 1,
                      inits = initsfunction)
```


## JAGS output for the MLR model

```{r message = FALSE, warning = FALSE}
summary(posterior_MLR)
```

```{r fig.align = "center", message = FALSE}
plot(posterior_MLR, vars = "beta1")
```

```{r fig.align = "center", message = FALSE}
plot(posterior_MLR, vars = "beta2")
```

```{r fig.align = "center", message = FALSE}
plot(posterior_MLR, vars = "sigma")
```

```{r fig.align = "center", message = FALSE}
post <- as.mcmc(posterior_MLR)
post %>% as.data.frame %>%
  gather(parameter, value) -> post2
ggplot(post2, aes(value)) +
  geom_density() + facet_wrap(~ parameter, ncol = 3) +
  theme(strip.text.x = element_text(size=8))
```


```{r fig.align = "center", message = FALSE, echo = FALSE}
# Add priors to graph

m0 <- the_data$mu0
m1 <- the_data$mu1
m2 <- the_data$mu2
m3 <- the_data$mu3
m4 <- the_data$mu4
m5 <- the_data$mu5
m6 <- the_data$mu6
m7 <- the_data$mu7

s0 <- 1 / sqrt(the_data$g0)
s1 <- 1 / sqrt(the_data$g1)
s2 <- 1 / sqrt(the_data$g2)
s3 <- 1 / sqrt(the_data$g3)
s4 <- 1 / sqrt(the_data$g4)
s5 <- 1 / sqrt(the_data$g5)
s6 <- 1 / sqrt(the_data$g6)
s7 <- 1 / sqrt(the_data$g7)

x <- seq(m0 - 3 * s0, m0 + 3 * s0, length.out = 100)
prior1 <- data.frame(value = x,
                     density = dnorm(x, m0, s0),
                     parameter = "beta0")
x <- seq(m1 - 3 * s1, m1 + 3 * s1, length.out = 100)
prior2 <- data.frame(value = x,
                     density = dnorm(x, m1, s1),
                     parameter = "beta1")
x <- seq(m2 - 3 * s2, m2 + 3 * s2, length.out = 100)
prior3 <- data.frame(value = x,
                     density = dnorm(x, m2, s2),
                     parameter = "beta2")
x <- seq(m3 - 3 * s3, m3 + 3 * s3, length.out = 100)
prior4 <- data.frame(value = x,
                     density = dnorm(x, m3, s3),
                     parameter = "beta3")
x <- seq(m4 - 3 * s4, m4 + 3 * s4, length.out = 100)
prior5 <- data.frame(value = x,
                     density = dnorm(x, m4, s4),
                     parameter = "beta4")
x <- seq(m5 - 3 * s5, m5 + 3 * s5, length.out = 100)
prior5 <- data.frame(value = x,
                     density = dnorm(x, m5, s5),
                     parameter = "beta5")
x <- seq(m6 - 3 * s6, m6 + 3 * s6, length.out = 100)
prior6 <- data.frame(value = x,
                     density = dnorm(x, m6, s6),
                     parameter = "beta6")
x <- seq(m7 - 3 * s7, m7 + 3 * s7, length.out = 100)
prior7 <- data.frame(value = x,
                     density = dnorm(x, m7, s7),
                     parameter = "beta7")
prior8 <- data.frame(value = x,
                     density = dgamma(1 / x ^ 2,
                                      shape = the_data$a,
                                      rate = the_data$b) * (2 / x ^ 3),
                     parameter = "sigma")

prior <- rbind(prior1, prior2, prior3, prior4, prior5, prior6, prior7, prior8)
prior$Type <- "Prior"
post2$Type <- "Posterior"

ggplot(post2, aes(value)) +
  geom_density() +
  geom_line(data = prior, aes(value, density),
            linetype = 2) +
  facet_wrap(~ parameter, ncol = 3) +
  theme(strip.text.x = element_text(size=18))+
  xlim(-0.5, 1.2) +
  theme_grey(base_size = 18, base_family = "")
```


