---
title: "Bayesian Inference for a Mean (R scripts)"
author: "Jingchen (Monika) Hu"
date: "MATH 347 Bayesian Statistics"
output:
  pdf_document: default
  html_document:
    number_sections: yes
institute: Vassar College
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
require(gridExtra)
require(ProbBayes)
require(tidyverse)
crcblue <- "#2905a1"
knitr::opts_chunk$set(echo = TRUE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

# Installing the necessary packages

```{r, eval = FALSE}
install.packages("devtools")
require(devtools)
devtools::install_github("bayesball/ProbBayes")

require(ggplot2)
require(gridExtra)
require(ProbBayes)
require(tidyverse)
crcblue <- "#2905a1"
```

# Example: Expenditures in the Consumer Expenditure Surveys

## The TOTEXPPQ variable

```{r message = FALSE}
CEsample = read_csv("CEsample1.csv")

summary(CEsample$TotalExpLastQ)
sd(CEsample$TotalExpLastQ)
```

## The TOTEXPPQ variable cont'd

```{r, fig.align = "center"}
ggplot(data = CEsample, aes(TotalExpLastQ)) +
  geom_density(color = crcblue, size = 1) + 
  labs(title = "Total expenditure last Q") +
  theme_grey(base_size = 8, base_family = "") 
```


## Log transformation of the TOTEXPPQ variable

```{r, fig.align = "center"}
CEsample$LogTotalExpLastQ <- log(CEsample$TotalExpLastQ)
ggplot(data = CEsample, aes(LogTotalExpLastQ)) +
  geom_density(color = crcblue, size = 1) + 
  labs(title = "Total expenditure last Q (log)") +
  theme_grey(base_size = 8, base_family = "") 
```


## Simulate posterior draws of $\mu$

```{r}
mu_0 <- 5
sigma_0 <- 1
phi_0 <- 1/sigma_0^2
ybar <- mean(CEsample$LogTotalExpLastQ)
phi <- 1.25
n <- dim(CEsample)[1]
mu_n <- (phi_0*mu_0+n*ybar*phi)/(phi_0+n*phi)
sd_n <- sqrt(1/(phi_0+n*phi))

set.seed(123)
S <- 1000
mu_post <- rnorm(S, mean = mu_n, sd = sd_n)
df <- as.data.frame(mu_post)
```

## Simulate posterior draws of $\mu$ cont'd

```{r, fig.align = "center"}
ggplot(data = df, aes(mu_post)) + 
  geom_density(color = crcblue, size = 1) + 
  labs(title = "Posterior density") +
  xlab(expression(mu)) +
  theme_grey(base_size = 8, base_family = "") 
```


## Simulate posterior draws of $\sigma$

```{r}
alpha <- 1
beta <- 1
mu <- 8
n <- dim(CEsample)[1]
alpha_n <- alpha+n/2
beta_n <- beta+1/2*sum((CEsample$LogTotalExpLastQ-mu)^2)

set.seed(123)
S <- 1000
invsigma2_post <- rgamma(S, shape=alpha_n, rate=beta_n)
df <- as.data.frame(invsigma2_post)
```


## Simulate posterior draws of $\sigma$ cont'd

```{r, fig.align = "center"}
ggplot(data = df, aes(invsigma2_post)) + 
  geom_density(color = crcblue, size = 1) + 
  labs(title = "Posterior density") +
  xlab(expression(1/sigma^2)) + 
  theme_grey(base_size = 8, base_family = "") 
```


# Bayesian inference for unknown mean $\mu$ (Lab 2)
   
# Recap
